A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-01-17 BSS7782.tex
updated tex file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Compact Binaries of BS Twins from Stellar Triples}
\shortauthors{Leigh \& Portegies Zwart}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

We propose a formation mechanism for twin blue stragglers in compact
binaries that involve mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
this scenario to the observed double blue straggler system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin blue
stragglers: (1) For the outer tertiary orbit, the initial orbital
period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$
1100 days, assuming initial masses for the inner binary components of
$m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After
Roche-lobe overflow, the outer star turns into a white dwarf (WD) of
0.43 to 0.54\,\MSun. There is a correlation between the mass of this
WD and the outer orbital period in the sense that more massive white
dwarfs will be on wider orbits.  (3) The rotational axes of both the
BSs in the inner binary will be aligned with each other and the
orbital plane of the outer tertiary WD. (4) The BSs in the inner
binary will have roughly the same masses, independent of their initial
masses.  This follows from the more copious accretion of mass onto the
initially lower mass star.  This star should, therefore, be enriched
more effectively by the accreted material.  As a result, one of the
blue stragglers will appear to be more enriched by He, C and O
if the donor started to overflow its Roche lobe at the red
giant branch, or with s-process elements, if the donor was on the
asymptotic giant branch.  (5) Relative to old dense clusters with high-velocity dispersions, twin BSs in close binaries formed from the
proposed mechanism should be more frequent in the Galactic field and
younger open clusters with ages $\lesssim$ 4-6 Gyr, since then the
donor will have a radiative envelope. (6) the orbit of the binary BS
will be close (typically $\aplt 0.3$\,au) and close to circular ($e
\aplt 0.2$).

@@

Blue straggler (BS) stars are brighter and bluer than the
main-sequence turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,2014ApJ...782...49S}.  Two primary
channels for BS formation have been proposed: mass transfer from an
evolved donor on to a main-sequence star in a binary star system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving main-sequence stars possibly
mediated via binaries
\citep[e.g.][]{1975AJ.....80..809H,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H,2018arXiv181100058P}.
Other possible, albeit related, formation mechanism includes mergers
of close main-sequence binaries (MS-MS) \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, many BSs
exist with observed properties that defy these simple scenarios.  For
example, in the old open cluster M67, there lurks a candidate triple
system that is posited to host two BSs
\citep{2001A&A...375..375V,2003AJ....125..810S}.  The observations
suggest that the outer tertiary is itself a BS, with a mass $\sim$ 1.7
M$_{\odot}$ and orbiting the inner binary with a period of $\sim
1188.5$ days \citep{2003AJ....125..810S}.  The inner binary has a
period of only $\sim 1.068$ days \citep{2001A&A...375..375V}, and
hosts a BS of mass $\sim 2.52$ M$_{\odot}$.  In order to form this
system we require at least five stars in order to conserve the total
mass in the system \citep{2011MNRAS.410.2370L}.  As explained in
\citet{2011MNRAS.410.2370L}, this is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collision events is the most probable explanation for its origin
(i.e., a single interaction involving two multiples with two or more
stars is a more likely scenario to produce this system than two
back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G}.

Even more curious, there exists in the old open cluster NGC 188 a
double BS binary, called Binary 7782.  More generally, the BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for Binary 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass blue
stragglers.  During a given binary-binary interaction, the probability
that not one but two direct (MS-MS) collisions will occur is less than
$10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \cite{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011Sci...334.1380F}.  So, how did Binary 7782 form?

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner binary composed of two main-sequence
stars.  In section~\ref{sect:dyn}, we constrain the range of initial
(i.e., pre-mass transfer) orbital parameters for a hypothetical outer
tertiary companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process triple
system with orbital parameters that, according to our expectations,
for promising progenitors for the twin BS 7782.  The calculations are
performed using the Astrophysical Multipurpose Software Environment
(\texttt{AMUSE} for short) \cite{AMUSE} with a combination of stellar
evolution, hydrodynamical and gravitational simulations.  with these
calculations we further constrain the possible range of initial
parameters that naturally lead to twin BS with orbital parameters
similar to the 7782 system.  We summarize and discuss the implications
of our results for compact double BS binaries and, more generally,
mass transfer in stellar triples in Section~\ref{sect:discussion}.

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For clarity we now
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination.  These assumptions are also
supported by the population of observed low-mass triples
\cite{2010yCat..73890925T,moe18}.  This initial configuration for our
assumed formation scenario for Binary 7782, described below, is
depicted in figure~\ref{fig:fig1}.

@@

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it leaves the asymptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamically stable, for which we adopt eq.\,1 in
\cite{1999ASIC..522..385M,}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk
\citep{2002apa..book.....F}.  Using conservation of angular momentum,
we equate the specific angular momentum of the accreted mass at the
inner Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the centre of mass corresponding to the tertiary defined by the
Roche lobe is given by Equation 2 in \citep{1983ApJ...268..368E}.
Combining Equation 2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
Equation~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2} =
0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3} =
1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of Figure~\ref{fig:fig2}), an outer tertiary that is Roche
lobe-filling, and a dynamically hard outer tertiary orbit.  \simon{Why
  do you consider it important to mention that the orbit is hard, and
  for what environment did you consider this?}  Note that the range of
plotted orbital periods P$_{\rm in}$ corresponding to a contact state
for the inner binary lies outside the range of plotted values for
$P_{\rm in}$ (for components with radii of roughly 1 R$_{\odot}$),
since it does not contribute to constraining the outer orbital
properties).  The thick horizontal solid red line shows the allowed
range of outer semi-major axes, after folding in all of 
the aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for
which the tertiary star overfills its Roche lobe, the triple is (at
least initially) dynamically stable and the accretion stream forms a
gaseous ring around the inner binary.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782
  before Roche-lobe overflow.  The solid diagonal black line shows the
  period corresponding to the circularization radius a$_{\rm c}$ for
  the mass transfer stream coming from the outer star (i.e., at the
  onset of mass transfer), assuming a mass ratio of q $=$ 0.7 and
  component masses of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1
  M$_{\odot}$ and m$_{\rm 3} =$ 1.4 M$_{\odot}$.  We assume initial
  component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm
    2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and
  m$_{\rm 3}$ is computed for the outer tertiary according to our
  assumed mass ratio (with our fiducial case corresponding to q $=$
  0.7).  We assume completely conservative mass transfer for this
  exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm
    \odot}$ once it has become a WD.  The dashed diagonal black line
  shows a rough criterion for dynamical stability in the triple,
  approximately following \citet{mardling99} (i.e., a$_{\rm in} <$
  0.1a$_{\rm out}$ is required for long-term dynamical stability in
  equal-mass co-planar triples).  The vertical solid red lines show
  the outer orbital periods corresponding to the hard-soft boundary
  assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km
  s$^{-1}$.  The vertical dashed black lines show the maximum outer
  orbital period P$_{\rm out}$ for which the outer tertiary companion
  is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$
  200 R$_{\odot}$ (which corresponds to the maximum stellar radius
  reached on the AGB for the range of tertiary masses of interest to
  us.  The horizontal dashed red line shows the observed orbital
  period for Binary 7782, using its observed orbital period and our
  assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2}
  =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line
  shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  
\label{fig:fig2}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  Here we adopted
    a donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, donor curve is
    quite similar (see dotted dark-blue curve).  The red squares in
    the curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin blue straggler 7782.  The initial triple in which it
    possibly formed must at least have been dynamically stable. The
    minimal orbital separation for the inner binary for which the
    triple is stable is given by the lower green coloured curve.
    Donors which are smaller than about 100\,\RSun\, (dark blue curve
    indicated with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertical dotted line.  The
    orbital separation at which the donor star overfills its Roche
    lobe is indicated with the light-blue curve. The top curve (brown)
    shows an estimate of the final orbit separation of the outer star,
    and therefore of the final orbit of the white dwarf around the
    inner twin blue straggler.  For core masses $\apgt 0.5$\,\MSun\,
    the final orbital separation, after mass transfer, is smaller
    than the initial orbit.  Here we adopted an initial inner binary
    mass of (1.0+0.9)\,\MSun\, and a final twin-blue straggler mass of
    (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

When adopting a mass of the tertiary star of $m_3 = 1.4$\,\MSun, we
can constrain the initial parameters for the inner binary as well as
the orbit of the outer star after mass transfer. We first calculate
the stellar radius as a function of core mass. In
figure\,\ref{fig:tertiarymass_vs_size} we present this relation
calculated using the {\tt SeBa} stellar evolution code
\cite{1996A&A...309..179P} (dark blue curve).  The interruption in
this curve, around a core mass of $m_{\rm core} \sim 0.5$\,\MSun\, is
a result of the evolution along the horizontal branch, where the core
of the star continues to grow but the radius actually shrinks.
Roche-lobe overflow in this phase is not expected to happen, because
it would already have happened in an earlier evolutionary state of the
donor star, when it was bigger.

Adopting masses for the inner binary $m_1=1.1$\,\MSun\, and
$m_2=0.9$\,\MSun\, we can calculate the outer orbital separation at
the onset of Roche-lobe overflow $a_{\rm out}$, and subsequently the
maximum orbital separation for the inner binary for which the orbit is
stable and a circum-binary disk can form. These two limits are
presented as the light-blue and light green curves in
figure\,\ref{fig:tertiarymass_vs_size}.  The alotted region of
parameter space is then above the dashed horizontal line and to the
right of the vertical dotted line.

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after the phase of
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer $m_3$ and $m'_3$,
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a$) and after mass transfer
($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides an
estimate of the orbital separation of the white dwarf around the twin
blue straggler 7782. Tidal effects during the episode of mass tranfer
have probably circularized the orbit, althoug some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induces a small $e \ll 0.1$ eccentricity.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Compact Binaries of BS Twins from Stellar Triples}
\shortauthors{Portegies Zwart \& Leigh}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\author{Simon Portegies Zwart}
%\affiliation{Leiden Observatory \\
%Leiden University \\
%PO Box 9513, 2300 RA \\
%Leiden, the Netherlands}

@@

We propose a formation mechanism for twin blue stragglers (BSs) in compact
binaries that involves mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
this scenario to the observed double BS system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin BSs: (1) For the outer tertiary orbit, the initial orbital
period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$
1100 days, assuming initial masses for the inner binary components of
$m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After
Roche-lobe overflow, the outer star turns into a white dwarf (WD) of
0.43 to 0.54\,\MSun. There is a correlation between the mass of this
WD and the outer orbital period: more massive WDs will be on wider orbits.  (3) The rotational axes of both 
BSs will be aligned with each other and the
orbital plane of the outer tertiary WD. (4) The BSs 
will have roughly equal masses, independent of their initial
masses (since the initially lower mass star accretes the most).  The dominant accretor should, therefore, be enriched
more effectively by the accreted material.  As a result, one of the
BSs will appear to be more enriched by either He, C and O or by s-process elements, depending on 
if the donor started to overflow its Roche lobe on, respectively, the red
giant or asymptotic giant branch.  (5) Relative to old dense clusters with high-velocity dispersions, twin BSs in close binaries formed from the
proposed mechanism should be more frequent in the Galactic field and
younger open clusters with ages $\lesssim$ 4-6 Gyr, since then the
donor will have a radiative envelope. (6) the orbit of the binary BS
will have a small semi-major axis (typically $\aplt 0.3$\,au) and be close to circular ($e
\aplt 0.2$).

@@

Blue straggler stars are brighter and bluer than the
main-sequence (MS) turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,2014ApJ...782...49S}.  Two primary
channels for BS formation have been proposed: mass transfer from an
evolved donor on to a MS star in a binary star system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving MS stars possibly
mediated via binaries
\citep[e.g.][]{1975AJ.....80..809H,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H,2018arXiv181100058P}.
Other possible, albeit related, formation mechanisms include mergers
of close MS-MS binaries \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, many BSs
exist with observed properties that defy these simple scenarios.  For
example, in the old open cluster (OC) M67, there lurks a candidate triple
system that is posited to host two BSs
\citep{2001A&A...375..375V,2003AJ....125..810S}.  The observations
suggest that the outer tertiary is itself a BS, with a mass $\sim$ 1.7
M$_{\odot}$ and orbiting the inner binary with a period of $\sim
1188.5$ days \citep{2003AJ....125..810S}.  The inner binary has a
period of only $\sim 1.068$ days \citep{2001A&A...375..375V}, and
hosts a BS of mass $\sim 2.52$ M$_{\odot}$.  In order to reproduce the total system mass 
we require at least five stars \citep{2011MNRAS.410.2370L}.  This is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collisions is the most probable explanation for its origin
(instead of back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G,2011MNRAS.410.2370L}.

Even more curious, there exists in the old OC NGC 188 a
double BS binary, called Binary 7782.  The BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for Binary 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass BSs.  During a given binary-binary interaction, the probability
that not one but two direct (MS-MS) collisions will occur is less than
$10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \cite{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011Sci...334.1380F}.  So, how did Binary 7782 form?

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner binary composed of two MS 
stars.  In section~\ref{sect:dyn}, we constrain the range of initial
(i.e., pre-mass transfer) orbital parameters for a hypothetical outer
tertiary companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process in this triple
system. We adopt orbital parameters that, according to our expectations,
are most promising for the progenitors of the twin BS 7782.  The calculations are
performed using the Astrophysical Multipurpose Software Environment
(\texttt{AMUSE} for short) \cite{AMUSE} with a combination of stellar
evolution, hydrodynamical and gravitational simulations.  with these
calculations we further constrain the possible range of initial
parameters that naturally lead to twin BSs with orbital parameters
similar to the 7782 system.  We summarize and discuss the implications
of our results for compact double BS binaries and, more generally,
mass transfer in stellar triples in Section~\ref{sect:discussion}.

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited by a tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For clarity we 
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination.  These assumptions are also
supported by the population of observed low-mass triples
\cite{2010yCat..73890925T,moe18}.  This initial configuration for our
assumed formation scenario for Binary 7782, described below, is
depicted in figure~\ref{fig:fig1}.

@@

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it leaves the asymptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamically stable, for which we adopt eq.\,1 in
\cite{1999ASIC..522..385M,}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk
\citep{2002apa..book.....F}.  Using conservation of angular momentum,
we equate the specific angular momentum of the accreted mass at the
inner Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the centre of mass corresponding to the tertiary defined by the
Roche lobe is given by eq.\,2 in \citep{1983ApJ...268..368E}.
Combining eq.\,2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
eq.~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2} =
0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3} =
1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of figure~\ref{fig:fig2}), and the assumption of an outer tertiary that is Roche
lobe-filling.  
%\simon{Why do you consider it important to mention that the orbit is hard, and for what environment did you consider this? - NL: Deleted.}  
Note that the range of
plotted orbital periods P$_{\rm in}$ corresponding to a contact state
for the inner binary lies outside the range of plotted values for
$P_{\rm in}$ (for components with radii of 1 R$_{\odot}$),
since it does not contribute to constraining the outer orbital
properties.  The thick horizontal solid red line shows the allowed
range of outer semi-major axes, after folding in all of 
the aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days.
%, for
%which the tertiary star overfills its Roche lobe, the triple is (at
%least initially) dynamically stable and the accretion stream forms a
%gaseous ring around the inner binary.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782
  before Roche-lobe overflow.  The solid diagonal black line shows the
  period corresponding to the circularization radius a$_{\rm c}$ for
  the mass transfer stream coming from the outer star (i.e., at the
  onset of mass transfer), assuming a mass ratio of q $=$ 0.7 and
  component masses of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1
  M$_{\odot}$ and m$_{\rm 3} =$ 1.4 M$_{\odot}$.  We assume initial
  component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm
    2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and
  m$_{\rm 3}$ is computed for the outer tertiary according to our
  assumed mass ratio (with our fiducial case corresponding to q $=$
  0.7).  We assume completely conservative mass transfer for this
  exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm
    \odot}$ once it has become a WD.  The dashed diagonal black line
  shows a rough criterion for dynamical stability in the triple,
  approximately following \citet{mardling99} (i.e., a$_{\rm in} <$
  0.1a$_{\rm out}$ is required for long-term dynamical stability in
  equal-mass co-planar triples).  The vertical solid red lines show
  the outer orbital periods corresponding to the hard-soft boundary
  assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km
  s$^{-1}$.  The vertical dashed black lines show the maximum outer
  orbital period P$_{\rm out}$ for which the outer tertiary companion
  is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$
  200 R$_{\odot}$ (which corresponds to the maximum stellar radius
  reached on the AGB for the range of tertiary masses of interest to
  us; see figure~\ref{fig:tertiarymass_vs_size}).  The horizontal dashed red line shows the observed orbital
  period for Binary 7782, using its observed orbital period and our
  assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2}
  =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line
  shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  
\label{fig:fig2}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  Here we adopt
    a donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, donor the curve is
    quite similar (see dotted dark-blue curve).  The red squares in
    the curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin BS 7782.  The initial triple in which it
    possibly formed must at least have been dynamically stable. The
    minimal orbital separation for the inner binary for which the
    triple is stable is given by the lower green coloured curve.
    Donors which are smaller than about 100\,\RSun\, (dark blue curve
    indicated with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertical dotted line.  The
    orbital separation at which the donor star overfills its Roche
    lobe is indicated with the light-blue curve. The top curve (brown)
    shows an estimate of the final orbital separation of the outer star,
    and therefore of the final orbit of the WD around the
    inner twin BSs.  For core masses $\apgt 0.5$\,\MSun\,
    the final orbital separation, after mass transfer, is smaller
    than the initial orbit.  Here we adopted an initial inner binary
    mass of (1.0+0.9)\,\MSun\, and a final twin BS mass of
    (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

Adopting a mass for the tertiary star of $m_3 = 1.4$\,\MSun, we
can constrain the initial parameters for the inner binary as well as
the orbit of the outer star after mass transfer. We first calculate
the stellar radius as a function of core mass. In
figure\,\ref{fig:tertiarymass_vs_size} we present this relation
calculated using the {\tt SeBa} stellar evolution code
\cite{1996A&A...309..179P} (dark blue curve).  The interruption in
this curve, around a core mass of $m_{\rm core} \sim 0.5$\,\MSun\, is
a result of the evolution along the horizontal branch, where the core
of the star continues to grow but the radius actually shrinks.
Roche-lobe overflow in this phase is not expected to happen, because
it would already have happened in an earlier evolutionary state of the
donor star, when it was bigger.

Adopting masses for the inner binary $m_1=1.1$\,\MSun\, and
$m_2=0.9$\,\MSun\, we can calculate the outer orbital separation at
the onset of Roche-lobe overflow $a_{\rm out}$, and subsequently the
maximum orbital separation for the inner binary for which the orbit is
stable and a circumbinary disk can form. These two limits are
presented as the light blue and light green curves in
figure\,\ref{fig:tertiarymass_vs_size}.  The alotted region of
parameter space is then above the dashed horizontal line and to the
right of the vertical dotted line.

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after 
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer $m_3$ and $m'_3$,
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a$) and after mass transfer
($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides a prediction for 
the current orbital separation of the WD around the twin
BS 7782. Tidal effects during mass tranfer
have probably circularized the orbit, although some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induces a small $e \ll 0.1$ eccentricity.

Having limited parameter space for the formation of the twin BS 7782, we continue by performing a series of simulations to
investigate the accretion and changes to the inner orbits of triple
systems in this range of parameters.
